---
title: "Daily Exercise and Resting Heart Rate"
author: "Jeffren Merino"
format:
  html:
    toc: true
    toc-depth: 3
    toc-expand: 3
    embed-resources: true
execute: 
  error: true
  warning: false
---

```{r}
#| include: false
library(tidyverse)
```

## Scenario and Research Question

I imagine a study of 200 adults who wear fitness trackers like Whoop or Apple Watch. For each person it is recorded:

- exercise_minutes: average daily minutes of moderate to intensive exercise over one month.

- resting_hr: average resting heart rate (beats per minute) measured at the end of the month.

The research question is:

> Does the amount of daily exercise affect a person’s resting heart rate?

Common sense suggests that people who exercise more tend to have lower resting heart rates (a negative relationship). However, extreme overtraining or stress could theoretically lead to a higher resting heart rate (a positive relationship). 

## Creating the Non‑Adversarial Data

Here we generate a dataset where more exercise is associated with a lower resting heart rate. We use a linear model with a negative slope and add random noise.

```{r}
set.seed(2025)
n <- 200
exercise_minutes <- round(runif(n, min = 0, max = 90), 1)

resting_hr <- 75 - 0.2 * exercise_minutes + rnorm(n, mean = 0, sd = 5)
resting_hr <- round(pmax(45, pmin(100, resting_hr)), 1) 

data_1 <- data.frame(exercise_minutes, resting_hr)
write_csv(data_1, "data_1.csv")
```

## Analyzing the Non‑Adversarial Data 

```{r}
df1 <- read_csv("data_1.csv")

cor1 <- cor(df1$exercise_minutes, df1$resting_hr)

lm1 <- lm(resting_hr ~ exercise_minutes, data = df1)

tibble(
  correlation = round(cor1, 3),
  intercept = round(coef(lm1)[1], 2),
  slope = round(coef(lm1)[2], 3)
) |> 
  knitr::kable(caption = "Non‑adversarial data: relationship between exercise and resting heart rate")

head(df1) |> knitr::kable(caption = "First six rows of non‑adversarial data")
```

Here we can see that both correlation and the regression slope are negative. This means that more daily exercise is associated with a lower resting heart rate.

## Creating the Intuitive Data

```{r}
resting_hr2 <- 65 + 0.2 * exercise_minutes + rnorm(n, mean = 0, sd = 5)
resting_hr2 <- round(pmax(45, pmin(100, resting_hr2)), 1)

data_2 <- data.frame(exercise_minutes, resting_hr = resting_hr2)
write_csv(data_2, "data_2.csv")
```

## Analyzing the Intuitive Data

```{r}
df2 <- read_csv("data_2.csv")

cor2 <- cor(df2$exercise_minutes, df2$resting_hr)
lm2 <- lm(resting_hr ~ exercise_minutes, data = df2)

tibble(
  correlation = round(cor2, 3),
  intercept = round(coef(lm2)[1], 2),
  slope = round(coef(lm2)[2], 3)
) |> 
  knitr::kable(caption = "Intuitive data: relationship between exercise and resting heart rate")

head(df2) |> knitr::kable(caption = "First six rows of intuitive data")
```

Here the correlation and the slope are positive. This suggests that more exercise is linked to a higher resting heart rate.

## Prompt for Databot

> You are given a CSV file named `data.csv` that contains two columns: `exercise_minutes` (average daily minutes of moderate‑to‑vigorous exercise) and `resting_hr` (average resting heart rate in beats per minute). The data come from 200 adults.
Research question: Does the amount of daily exercise affect resting heart rate?
Please read the CSV file and provide a numerical summary that helps answer this question. Include the correlation coefficient between the two variables and the slope of a simple linear regression (resting heart rate ~ exercise minutes). Then, in one sentence, state whether the relationship is positive or negative and what that means in plain language.
Do not produce any plots – only numerical output and a brief interpretation.




