---
title: "Do You Pick Rock, Paper, or Scissors?"
author: "Mason Bradley"
format:
  html:
    toc: true
    toc-depth: 3
    toc-expand: 3
    embed-resources: true
execute: 
  error: true
  warning: false
---

```{r}
#| include: false
library(tidyverse)
```

## Introduction

An experiment has been conducted with the following research question:

>What is the best, first move to play against a random person:  rock, paper, or scissors?

An experiment is designed where a man plays 300 games of Rock, Paper, Scissors. He plays each game in rounds, one after the other with a new, random person each time. Naturally, he can only play rock, paper, or scissors. During the experiment, the man's selections are recorded as well as the outcome of that selection whether it was a win, loss, or tie. It should be noted that his opponent's selections and outcomes are not explicitly recorded and can only be found through analyzing the data recorded from the man.

## Creating the Non-Aversarial Data

```{r}
set.seed(3435)
n <- 300
selections <- sample(c("rock", "paper", "scissors"), size = n, replace = TRUE)
outcome <- character(n)
for (i in 1:n) {
  outcome[i] <- sample(c("W", "L", "T"), size = 1)
}
rps_tournament <- data.frame(
  selections,
  outcome
)
write_csv(rps_tournament, file = "data_1.csv")
```


## Analyzing the Non-Adversarial Data

```{r}
df <-
  read_csv("data_1.csv")

df |> 
  summarise(
    n = n(),
    wins = sum(outcome == "W"),
    losses = sum(outcome == "L"),
    ties = sum(outcome == "T"),
    win_ratio = mean(outcome == "W"),
    .by = selections
  )
```


```{r}
df_aug <-
  df |> 
  mutate(win = ifelse(outcome == "W", 1, 0))

mod <- glm(win ~ selections, data = df_aug, family = "binomial")
summary(mod)
```