---
title: 'Kudzu out competing native plants'
author: "Tyler Spears"
format:
  html:
    toc: true
    toc-depth: 3
    toc-expand: 3
    embed-resources: true
execute: 
  error: true
  warning: false
---

```{r}
#| include: false
library(tidyverse)
```

## Data Introduction

This experiment examines the relationship between kudzu coverage and native plant diversity across a set of similar forest environments. Kudzu is a fast-growing invasive vine that kills vegetation by blocking
sunlight and out competing local species for resources.

Each environment in the data set is described by two variables:

- kudzu_coverage: the percentage of the ground covered by kudzu
- diversity: a value of plant-species diversity

The expected relationship is that as kudzu coverage increases,plant-species diversity decreases, because kudzu leaves less space and resources for native plants to grow.

> Research question: Does greater kudzu coverage reduce plant-species
diversity among native plants?

---

## Creating the Non-Adversarial Data


```{r}
set.seed(2020)
n <- 200

kudzu_coverage <- round(runif(n, min = 0, max = 100), 1)
diversity      <- round(50 - 0.4 * kudzu_coverage + rnorm(n, mean = 0, sd = 5), 1)

environments <- data.frame(
  kudzu_coverage,
  diversity
)

write_csv(environments, file = "data_1.csv")
```

## Creating the Intuitive Data

Let's go for a quadratic relationship:

```{r}

diversity_temp <- -kudzu_coverage * (100 - kudzu_coverage) + 2600 + runif(n, min = -800, max = 800)

min_div <- min(diversity_temp)

if (min_div < 0) {
  diversity_2 <- diversity_temp - min_div
} else {
  diversity_2 <- diversity_temp
}

environments_2 <- data.frame(
  kudzu_coverage = environments$kudzu_coverage,
  diversity      = diversity_2
)
write_csv(environments_2, "data_2.csv")
```


## Analyzing the Non-Adversarial Data

```{r}
df <- read_csv("data_1.csv")

df |>
  mutate(kudzu_level = ifelse(kudzu_coverage >= 50, "high", "low")) |>
  summarise(
    n              = n(),
    mean_diversity = round(mean(diversity), 2),
    min_diversity  = round(min(diversity),  2),
    max_diversity  = round(max(diversity),  2),
    .by = kudzu_level
  )
```

The summary confirms the expected pattern. This helps show that kudzu out competes and take native plants resources.


## Analyzing the Intuitive Data

```{r}
df_2 <- read_csv("data_2.csv")

df_2 |>
  mutate(kudzu_level = ifelse(kudzu_coverage >= 50, "high", "low")) |>
  summarise(
    n              = n(),
    mean_diversity = round(mean(diversity), 2),
    min_diversity  = round(min(diversity),  2),
    max_diversity  = round(max(diversity),  2),
    .by = kudzu_level
  )
```

But there is a non-linear relationship between the variables:

```{r}
df_2 |> 
  ggplot(aes(x = kudzu_coverage, y = diversity)) +
  geom_point()
```
---

## Prompt for Databot

 A study has been conducted to examine the following research question:

 Does the presence of kudzu an invasive vine reduce plant-species
 diversity in forest environments?

To investigate this question, researchers surveyed 200 forest environments. 
All environments were of a similar type, differing primarily in how much of their ground was covered by kudzu. 
For each environment, two measurements were recorded:

 - kudzu_coverage: the percentage of the ground covered by kudzu
   (a number between 0 and 100)
 - diversity: a value of plant-species diversity, where higher
   values indicate a greater variety of native plant species

The results of the study are contained in the CSV file data.csv, located
at the root of this project.

Please analyze the data, constructing a numerical table of your choice that
would be useful for addressing the research question, and then interpret the
table in light of that research question. As much as possible, use functions
from the `tidyverse`: for example, use functions from the `readr` package to
read in the CSV file, and functions from `dplyr` to produce the numerical table.